---
title: "ph1861_hw2_ygu5"
author: "Yue Gu"
date: "2023-11-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


Meta-Analysis Problem #1 (50 points)
The following paper reviewed and meta-analyzed the proportion of Campylobacter cases that develop chronic sequelae, based on studies published prior to July 2011. (Keithlin 2014, “Systematic review and meta-analysis of the proportion of Campylobacter cases that develop chronic sequelae”). Import the number of cases that developed reactive arthritis (ReA) from page 7 of the article and follow the steps below to conduct a meta-analysis of the proportion/prevalence. 

Please input data in the exact order from the paper.
Calculate the proportion of people with Campylobacter who developed a ReA for each study.
Add a count of 0.5 both to those reporting ReA outcome of 0% and to those totals.
Calculate the logit of the outcome and logit of the standard error using the equation from Lipsey & Wilson. [Eq 3.5, p40].

```{r}
# read data
keith_data = read.csv("./data/Keithlin2014_ReA.csv") %>% 
  janitor::clean_names() %>% 
  mutate(prop_seq = round(number_developing_sequelae/number_of_people_with_campylobacter,4)) %>% 
  mutate(number_of_people_with_campylobacter_add = ifelse(prop_seq==0.0000, number_of_people_with_campylobacter+0.5, number_of_people_with_campylobacter),
         number_developing_sequelae_add = ifelse(prop_seq==0.0000, number_developing_sequelae+0.5, number_developing_sequelae)) %>% 
  mutate(prop_seq_add = number_developing_sequelae_add/number_of_people_with_campylobacter_add)

  
```



